Design, Implementation, and Evaluation of Open Power Quality

### 2020-05-21 02-intro.tex
checckpoint

One impact of adding renewable energy generation to an electrical grid is that maintaining adequate power quality is much more challenging. This problem is inversely proportional to the size of the electrical grid. For example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai.  (In contrast, the European electrical grid is approximately 600 GW and the U.S. continental grid is over 1000 GW.) For small grids, the unpredictable nature of power generation for distributed, intermittent renewables can quickly create problems for grid stability. In the case of Hawaii, consumer demand for grid-tied rooftop solar exceeded the ability of Hawaiian Electric to manage, resulting first in complex and expensive "interconnection studies" \cite{trabish_solar_2014,anastasi_energy_2009}, and later in the requirement for new installations to include batteries and grid-disconnected operation.

@@

These assumptions might not hold in grids with distributed, intermittent generation by end-users, such as is the case with roof top solar. In these cases, the unpredictable nature of generation can lead to local power quality problems that are not experienced by the substation, and that are not due to any single individual, but rather the bottom-up power generation architecture of the grid. \cite{nakafuji_back--basics_2011}

@@

{\em 4. Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers. For example, in Hawaii, utilities are required to make a "best effort" to provide non-harmful voltage and frequency values, but are only required to publicly report on power outages of more than 3 minutes. There is no requirement for utilities to report potentially harmful levels of voltage, frequency, THD, or transients as long as they do not lead to outages. Indeed, in places like Hawaii, there is not even infrastructure in place to enable utilities to collect that information, even if they were asked to report on it. For more details, see Hawaiian Electric Tariff Rule No. 2 Character of Service and Hawaiian Electric Tariff Rule No. 16 Interruption of Service.

Providing power quality data to address the above issues requires solving several difficult technical problems. First, monitoring below the substation level increases the number of required monitoring devices by two to three orders of magnitude. This has significant implications for the cost of monitoring in terms of the devices, and the sheer amount of low-level power quality data that will be produced. Second, this low-level data must be processed in a manner that does not require inordinate amounts of processing, network bandwidth, or storage. Finally, the data must lead to actionable information.

@@

The goal of the OPQ Project is to provide an unbiased, independent, third party source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation. OPQ accomplishes this through a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services. An OPQ sensor network is intended to be useful to consumers who want to better understand the performance of their public utilities, to researchers who wish to design improvements to grid management, and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.

{\em OPQ Box.} First, our power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75. This is from 10 to 100 times less expensive than commercial power quality monitors. This means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building. Similarly, instead of monitoring power quality at the substation level of the grid, OPQ Boxes make it feasible to deploy dozens or hundreds into a community to obtain fine-grained data about the impact of solar or other renewables as directly perceived by the customer.

{\em Cloud native.} Second, our architecture is "cloud native", which means that OPQ Boxes are not designed for "stand alone", autonomous use, unlike current commercially available power quality monitors. Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and analyzed. This creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network. It means our cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud. This enables OPQ installations to be simultaneously responsive and scalable.

{\em Extensibility.} Third, all OPQ software subsystems are designed for extensibility and interoperation. Our middleware components (Mauka and Makai) provide a plugin architecture. Our visualization component (View) is built from modular UI elements using React. We are also developing API endpoints for interoperation with other software systems. This means that OPQ is not a monolithic, closed system with a frozen feature set, but rather a design environment for experimentation with advanced classification and analysis of power quality data.

@@

Finally, while we believe OPQ offers a compelling combination of capabilities, we want to be clear that all designs involve trade-offs. As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs. In some cases, OPQ may not be the appropriate choice. In other cases, a hybrid solution consisting of OPQ in combination with other technologies for power monitoring may be most appropriate. Our goal in this paper is to describe OPQ well enough for organizations to make a thoughtful decision about the use of our technology.

>>>>

One impact of adding renewable energy generation to an electrical grid is that maintaining adequate power quality is much more challenging. This problem is inversely proportional to the size of the electrical grid. For example, each island in the State of Hawaii maintains its own electrical grid, ranging from a 1.8 GW grid for the island of Oahu to a 6 MW grid for the island of Molokai.  (In contrast, the size of the European electrical grid is approximately 600 GW and the size of the U.S. continental grid is over 1000 GW.) For small grids, the unpredictable nature of power generation by distributed, intermittent renewables can quickly create problems for grid stability. In the case of Hawaii, consumer demand for grid-tied rooftop solar exceeded the ability of Hawaiian Electric to manage, resulting first in complex and expensive "interconnection studies" \cite{trabish_solar_2014,anastasi_energy_2009}, and later in the requirement for new installations to include batteries and grid-disconnected operation.

@@

These assumptions might not hold in grids with distributed, intermittent generation by end-users, such as is the case with roof top solar. In these cases, the unpredictable nature of generation can lead to local power quality problems that are not experienced by the substation, and that are not due to any single individual, but rather the bottom-up power generation architecture of the grid. Nakafuji notes that Hawaiian Electric is contending with PV penetrations in excess of 60\% on certain distribution circuits, for which traditional rules of thumb for design of protection and distribution systems might not hold \cite{nakafuji_back--basics_2011}.

@@

{\em 4. Power quality data should be publicly available.} Electrical utilities are not required to be totally transparent about the quality of power they provide to consumers. For example, in Hawaii, utilities are required to make a "best effort" to provide non-harmful voltage and frequency values, but are only required to publicly report on power outages of more than 3 minutes. There is no requirement for utilities to report potentially harmful levels of voltage, frequency, THD, or transients as long as they do not lead to outages. Indeed, in places like Hawaii, there is not yet infrastructure in place to enable utilities to collect that information, even if they were asked to report on it.

Providing power quality data to address the above issues requires solving several difficult technical problems. First, monitoring below the substation level increases the number of required monitoring devices by two to three orders of magnitude. This has significant implications for the cost of monitoring in terms of the devices, and the sheer amount of low-level power quality data that will be produced. Second, this low-level data must be processed in a manner that does not require inordinate amounts of processing, network bandwidth, or storage. Finally, collecting the data is not useful if it cannot be converted into actionable information in a reasonable amount of time.

@@

The goal of the OPQ Project is to provide a scalable source of actionable power quality data about electrical grids, particularly those with high levels of distributed, intermittent power generation. OPQ accomplishes this through the design and implementation of a low-cost sensor network for power quality using custom power quality monitors with real-time, two way communication to a set of cloud-based services. An OPQ sensor network is designed to be useful to consumers who want to better understand the performance of their public utilities, to researchers who wish to design improvements to grid management, and to public policy makers responsible for designing and implementing regulatory frameworks for electrical utilities.  There are several key features of our project emerge from these goals.

{\em OPQ Box.} First, our custom power quality monitoring hardware device, called "OPQ Box" can be produced for approximately US\$75. This is from 10 to 100 times less expensive than commercial power quality monitors. This cost differential means, for example, that instead of monitoring power quality at the level of individual buildings, OPQ makes it economically feasible to monitor power quality on each floor or even in each room of a building. Similarly, instead of monitoring power quality at the substation level of the grid, OPQ Boxes make it feasible to deploy dozens or hundreds into a community to obtain fine-grained data about the impact of solar or other renewables as directly perceived by the customer.

{\em Cloud-native information architecture.} Second, our information architecture is "cloud native", which means that OPQ Boxes are not designed for stand-alone, autonomous use, unlike current commercially available power quality monitors. Instead, our hardware boxes and cloud-based services perform real-time, two-way communication over the Internet to determine what power quality data must be gathered and with what fidelity in order to most efficiently produce actionable information. Two-way communication between the OPQ Boxes and could services creates the ability to control individual OPQ Boxes based upon the global state of all OPQ Boxes in an OPQ sensor network. As a result, OPQ cloud services can obtain and analyze high fidelity wave form data only when desired, without the prohibitive network overhead of constantly communicating this data from boxes to the cloud. This enables OPQ installations to be simultaneously responsive and scalable.

{\em Plug-in analysis architecture.} Third, all OPQ software services are designed for extensibility and interoperation. Our middleware components (Mauka and Makai) provide a plugin architecture. Our visualization component (View) is built from modular UI elements using React.  This means that OPQ is not a monolithic, closed system with a frozen feature set, but rather a design environment for experimentation with advanced classification and analysis of power quality data.

@@

While we believe OPQ offers a compelling combination of capabilities, we want to be clear that all network designs involve trade-offs. As we will show below, other commercial and research solutions have features not provided by OPQ that may be important depending upon an organization's specific needs. In some cases, OPQ may not be the appropriate choice. In other cases, a hybrid solution consisting of OPQ in combination with other technologies for power monitoring may be most appropriate.

\subsection{Structure of this paper}

The next sections present an overview of the OPQ system architecture and information architecture. The following section presents related research and technology.  Section \ref{sec:makai} provides more detail on the OPQ Box, and Section \ref{sec:makai} provides more detail on our cloud services. Section \ref{sec:pilot} discusses the pilot deployment of a sensor network and its results. We conclude in Section \ref{sec:conclusion} with some proposals for future research.
